%------------------------------------------------------------------
% Dutch notes
%------------------------------------------------------------------
\documentclass[letterpaper,11pt]{article}
%---------------------
% Packages
%---------------------
\usepackage{color}
\usepackage[margin=1.0in]{geometry}
\usepackage{tabularx}
\usepackage{scrextend}
\usepackage{graphicx}
\usepackage{enumitem}
%---------------------

%---------------------
% Macros & Misc
%---------------------
\topmargin=-1.0in
\setlength{\textheight}{10in}
\definecolor{resgrey}{gray}{.85}
\newcommand{\sectionhead}[1]{{\large
        {\begin{minipage}[t]{\linewidth}{\textit{\textbf{#1
\vphantom{p\^{e}}}}\end{minipage}}}}}
\topmargin=-1.0in
\setlength{\textheight}{10in}

\pagenumbering{gobble}
\begin{document}
%-------------------------------------------------------------------------
% page 1
%-------------------------------------------------------------------------
%------------------------------------------------------------------
% Section: Heading
%------------------------------------------------------------------
\begin{flushleft}
    \LARGE{\textbf{Dutch Verbs}}
\end{flushleft}
\sectionhead{Introduction}
\par{In Dutch there are two main tenses, \textit{present} and \textit{past}.
These two tenses can be further divided into eight other tenses. Each of these
eight tenses require you to know both the correct conjugation and application.} 
\\
\begin{tabular}[t]{l l}
    \textit{Verb Tenses} \\
    \hline
    1: Simple Present  & 5: Future              \\
    2: Simple Past     & 6: Conditional         \\
    3: Present Perfect & 7: Future Perfect      \\
    4: Pluperfect      & 8: Conditional Perfect \\
    \hline
\end{tabular}
\\ \\
%------------------------------------------------------------------
% Section: Conjugation
%------------------------------------------------------------------
\sectionhead{Conjugation}
\par{In order to conjugate a verb correctly we must first identify the verb
    stem. Typically in Dutch verb infinitives end in the letters \textit{-en},
    so to get the correct verb stem we just subtract this ending. For example
    if we have the verb \textit{werken (to work)} the correct form of the verb
stem would be \textit{werk}.} 
\\
There are a few rules you must keep in mind when working with verb stems:
\begin{enumerate}[topsep=0pt, parsep=0pt, partopsep=0pt, itemsep=0pt]
    \item Long vowel infinitives require long vowel stems.
        \subitem \small{\textit{Maken} would have the stem \textit{Maak} to
        keep its long vowel sound.}
    \item A verb stem never ends in two identical consonants
        \subitem \small{\textit{Pakken} would have the stem
        \textit{Pak}.}
    \item A verb stem never ends in \textit{-v} or \textit{-z}, verbs with these
        endings instead are replaced with \textit{-f} or \textit{-s} respectively.
        \subitem \small{\textit{Leven} would have the stem \textit{Leef}.}
    \item The stem of an \textit{-ien} verb ends in \textit{-ie}.
        \subitem \small{\textit{Skien} would have the stem \textit{Skie}.}
\end{enumerate}
%------------------------------------------------------------------
% Section: Simple Present
%------------------------------------------------------------------
\sectionhead{Simple Present}
\par{The \textit{simple present} tense is used to refer to an event that takes place
right now, or the current state of an object.}
\\
\begin{tabular}[t]{l l}
    \textit{Conjugation} \\
    \hline
    \textit{Ik} + \textit{stem} + \textit{t} & \textit{We} + \textit{infinitive} \\
    \textit{Je} + \textit{stem} + \textit{t} & \textit{Jullie} + \textit{infinitive} \\
    \textit{Hij} + \textit{stem} + \textit{t} & \textit{Ze} + \textit{infinitive} \\
    \hline
\end{tabular}
\\ \\
In Dutch the simple present is used in four cases.
\begin{enumerate}[topsep=0pt, parsep=0pt, partopsep=0pt, itemsep=0pt]
    \item To refer to a momentary action that coincides wit the moment we are
    talking about it.
        \subitem \small{NL: Ik neem een hapje.}
        \subitem \small{EN: I am taking a bite.} 
    \item To refer to an ongoing, habitual, or repetitive action or state.
        \subitem \small{NL: Ze werkt bij de overheid.}
        \subitem \small{EN: She works for the government.}
    \item To refer to a future event in combination with an adverb of time.
        \subitem \small{NL: Ik neem er straks nog een.}
        \subitem \small{EN: I will have another one in a moment.}
    \item To refer to a hypothetical \textit{if-then} situation.  
        \subitem \small{NL: Als je te cola drinkt krijg je gaatjes in je tanden.} 
        \subitem \small{EN: If you drink to much coke you will get cavities in your teeth.} 
\end{enumerate}
%-------------------------------------------------------------------------
% page 2
%-------------------------------------------------------------------------
%-------------------------------------------------------------------------
%------------------------------------------------------------------
% Section: Simple Past
%------------------------------------------------------------------
\\ \\
\sectionhead{Simple Past}
\par{Before we go get into the simples past we must make a distinction between
    two types of verbs \textit{t-verbs} which have a crude stem ending of
    \textit{t,h,f,c,k,s,p} and \textit{d-verbs}. A good way to remember this
    rule is with the phrase \textit{pocket fish} which contains all consonant
    endings that classify a \textit{t-verb}.}
\\ 
\begin{tabular}[t]{l l}
    \textit{Conjugation t-verb} \\
    \hline
    \textit{Ik} + \textit{stem} + \textit{te} &  \textit{We} + \textit{stem} + \textit{ten} \\
    \textit{Je} + \textit{stem} + \textit{te} &  \textit{Jullie} + \textit{stem} + \textit{ten} \\
    \textit{Hij} + \textit{stem} + \textit{te} & \textit{Ze} + \textit{stem} + \textit{ten} \\
    \hline
\end{tabular}
\begin{tabular}[t]{l l}
    \textit{Conjugation d-verb} \\
    \hline
    \textit{Ik} + \textit{stem} + \textit{de} &  \textit{We} + \textit{stem} +
    \textit{den} \\
    \textit{Je} + \textit{stem} + \textit{de} &  \textit{Jullie} + \textit{stem}
    + \textit{den} \\
    \textit{Hij} + \textit{stem} +\textit{de} & \textit{Ze} + \textit{stem} +
    \textit{den} \\
    \hline
\end{tabular}
\\ \\
We use the \textit{simple past} tense in dutch when:
\begin{enumerate}[topsep=0pt, parsep=0pt, partopsep=0pt, itemsep=0pt]
    \item Past events that do not have any bearing on the present
        \begin{itemize}
            \item If the event or action is still relevant to the present time we
                generally use the present perfect. When referring to past events or
                actions, the present perfect is more common that the simple past.
                \subitem \small{NL: Karel de grote regeerde van 800 tot 814.} 
                \subitem \small{EN: Charlemagne reigned from 800 until 814.} 
        \end{itemize}
    \item Describing what went on during a certain past event.
        \begin{itemize}
            \item When we refer to an event that took place in the past, all information
                surrounding that event is set in the simple past.
                \subitem \small{NL: Tijdens de kabinetscrisis was de premier op
                vakantie. } 
                \subitem \small{EN: During the cabinet crisis, the prime-minister was on
                vacation.} 
        \end{itemize}
    \item When we introduce a past action or event by \textit{toen (when)}.
        \begin{itemize}
            \item When we point to a past event by using \textit{toen}, we
                generally use the simple past tense. If we use the perfect tense it all
                it must be the \textit{pluperfect tense}.
                \subitem \small{NL: Toen ik wakker werd, scheen de zon volop.} 
                \subitem \small{EN: When i woke up, the sun was shining brightly.} 
        \end{itemize}
\end{enumerate}
\\ \\
%------------------------------------------------------------------
% section: Present Perfect
%------------------------------------------------------------------
\sectionhead{Present Perfect}
\par{For the \textit{present perfect} tense, we generally use the verb
    \textit{hebben (to have)}, and \textit{zijn (to be)}. This can be a bit
    confusing because it is precisely what we said about the \textit{simple
    past}. There are no strict rules to tell you when to use which tense
    however. Generally speaking if you know when to use the \textit{simple past},
you can safely use the \textit{present perfect} in all other cases.}
\\
\begin{tabular}[t]{l l}
    \textit{Conjugation} \\
    \hline
    \textit{Ik} \textit{heb/ben} + \textit{p.p.} &  \textit{We hebben/zijn} +
    \textit{p.p.} \\
    \textit{Je} \textit{hebt/bent} + \textit{p.p.} & \textit{Jullie hebben/zijn}
    + \textit{p.p.} \\
    \textit{Hij} \textit{heeft/is} + \textit{p.p.} & \textit{Ze hebben/zijn} +
    \textit{p.p.} \\
    \hline
\end{tabular}
%------------------------------------------------------------------
% Section: Pluperfect
%------------------------------------------------------------------
\\ \\
\sectionhead{Pluperfect}
\par{The \textit{pluperfect} works in the same way as the \textit{present
perfect}, except that we use the \textit{simple past} tense of the verbs
\textit{hebben} and \textit{zijn}.} \\
\begin{tabular}[t]{l l}
    \textit{Conjugation} \\
    \hline
    \textit{Ik} + \textit{had/was} + \textit{p.p.} & \textit{We} + \textit{hadden/waren} + \textit{p.p.} \\
    \textit{Je} + \textit{had/was} + \textit{p.p.} & \textit{Jullie} + \textit{hadden/waren} + \textit{p.p.} \\
    \textit{Hij} + \textit{had/was} + \textit{p.p.} & \textit{Ze} + \textit{hadden/waren} + \textit{p.p.} \\
\end{tabular}
\\
\par{We use the pluperfect to refer to an event that occurred before another
past event.} 
\\
\begin{small}
    NL: Toen wij aankwamen, waren de meeste gasten al gearriveerd. \\
    EN: When we came, most guest had already arrived. \\
\end{small}
%------------------------------------------------------------------
% Section: Simple future
%------------------------------------------------------------------
\sectionhead{Simple Future}
\par{For the \textit{simple future} tense we use the auxiliary verb
    \textit{zullen (will)}. This tense however is quite a formal way to talk about
the future. Bellow we also show the more common way to refer to the future.} \\
\begin{tabular}[t]{l l}
    \textit{Conjugation} \\
    \hline
    Ik zal + \textit{infinitive} & We zullen + \textit{infinitive} \\
    Je zult/zal + \textit{infinitive} & Jullie zullen + \textit{infinitive} \\
    Hij zal + \textit{infinitive} & Ze zullen + \textit{infinitive} \\
    \hline
\end{tabular} \\
\begin{tiny}
    Note: Both zult and zal are equally correct.
\end{tiny}
\par{The \textit{zullen} + \textit{infinitive} construction is more similar to the English word \textit{shall} than to
\textit{will}. We use this form in the following situations.}
\begin{enumerate}[topsep=0pt, parsep=0pt, partopsep=0pt, itemsep=0pt]
    \item To express a promise or proposal.
        \subitem \small{NL: Ik zal het nooit meer doen!} 
        \subitem \small{EN: I shall not do it again.} 
    \item Stress that something will most certainly happen.
        \subitem \small{NL: Je zult dat nog nodig hebben.} 
        \subitem \small{EN: You are going to need it.} 
    \item Express that an event is likely going to take place.
        \subitem \small{NL: Hij zal het waarschijnlijk morgen bekendmaken.} 
        \subitem \small{EN: He will probably announce it tomorrow.} 
\end{enumerate}
\textit{\textbf{\small{Common Future}}}
\par{As previously mentioned the \textit{simple future} is a more formal
    grammatical construct. Instead the Dutch typically refer to the future using
what we'll call here the \textit{common future}.}
\\
Often in dutch the word \textit{gaan(to go)} is used to:
\begin{enumerate}[topsep=0pt, parsep=0pt, partopsep=0pt, itemsep=0pt]
    \item Express an intended action but no promise, proposal, or solemn plan.
        \subitem \small{NL: Ik ga vanavond pannenkoeken bakken.} 
        \subitem \small{EN: I am going to bake pancakes tonight.} 
    \item To say that an event is going to take place, without stressing the
        certainty or mentioning the probability.
        \subitem \small{NL: Zij gaat morgen een liedje voor ons zingen.} 
        \subitem \small{EN: Tomorrow, she is going to sing a song for us.} 
    \item If the point in time when the event is going to take place is explicitly
        mentioned, we often use the simple present.
        \subitem \small{NL: Ik bak vanavond pannenkoeken.} 
        \subitem \small{EN: Tonight, I will bake pancakes.} 
\end{enumerate}
%------------------------------------------------------------------
% Section: Future Perfect
%------------------------------------------------------------------
\sectionhead{Future Perfect}
\par{The future perfect is rather uncommon in Dutch. It is used when you want to
    say something with have been completed. For the conjugation of the
    \textit{future perfect}, you have to combine your knowledge of the
\textit{future} and \textit{perfect} tense.}
\\
\begin{tabular}[t]{l l}
    \textit{Conjunction} \\
    \hline
    Ik zal hebben/zijn + p.p. & We zullen hebben/zijn + p.p. \\
    Je zult/zal hebben/zijn + p.p. & Jullie zullen hebben/zijn + p.p. \\
    Hij zal hebben/zijn + p.p. & Ze zullen hebben/zijn + p.p. \\
    \hline
\end{tabular}
\\
%This is comparing two future tenses
\par{We use the \textit{future perfect} to say that something will have been
    completed at some point int he future. This tense is quite uncommon in Dutch
    instead they usually opt for the \textit{present perfect}. Bellow we will
compare these two forms and how they're used.}
\begin{enumerate}[topsep=0pt, parsep=0pt, partopsep=0pt, itemsep=0pt]
    \item \textit{Future Perfect - less common}
        \subitem \small{NL: Morgen zal ze het allemaal zijn vergeten.} 
        \subitem \small{EN: Tomorrow, she has forgotten all about it.} 
    \item \textit{Present Perfect - more common}
        \subitem \small{NL: Morgen is ze het allemaal vergeten.} 
        \subitem \small{EN: Tomorrow, she has forgotten all about it.} 
\end{enumerate}
%------------------------------------------------------------------
% Section: Conditional
%------------------------------------------------------------------
\sectionhead{Conditional}
\par{The \textit{conditional} tense is used to refer to hypothetical situations.
    The \textit{conditional} tense works the same as the future tense, however
    instead of the present tense of the irregular verb \textit{zullen} we use the
past tense.}
\\
\begin{tabular}[t]{l l}
    \textit{Conjugation} \\
    \hline
    \textit{Ik zou} + \textit{infinitive} & \textit{We zouden} + \textit{infinitive} \\
    \textit{Je zou} + \textit{infinitive} & \textit{Jullie zouden} + \textit{infinitive} \\
    \textit{Hij zou} + \textit{infinitive} & \textit{Ze zouden} + \textit{infinitive} \\
    \hline
\end{tabular}
\\ \\
%Sub-section: Will and would
\textbf{Will and would} \\
For more realistic, we can also use the future tense. If certain criteria are
met, then a certain hypothetical situation will be the case. In this case,
however the Dutch normally use the \textit{simple present}.
\\
%Sub-section: Present and future
\textbf{Present and future} \\
Although we use the past tense of the verb \textit{zullen}, we are referring to
hypothetical situations in the present or future. For past hypothetical
situations we use the conditional perfect.
\\
%Sub-section: If-then situations
\textbf{If-then situations} \\
For the verbs in the \textit{if-clause}, the English often use the
subjunctive (not \textit{he was} but \textit{her were}).
\\
\begin{small}
    \indent NL: Dat zou ik niet doen als ik jou was. \\
    \indent EN: I would not do that if I were you. \\
\end{small}
Sometimes, the \textit{if-then} structure is not clearly visible.
\\
\begin{small}
    \indent NL: We zouden ons maar vervelen. \textit{(als we er te lang bleven)} \\
    \indent EN: We would only be bored. \textit{(if we stayed there to long)} \\
\end{small}
%Sub-section: Supposed to be situations
\textbf{Supposed to be situations} \\
\textit{Zouden} is also used when we talk about what should be the case
according to our norms, plans, or expectations. \\
\begin{small}
    \indent NL: Hij zou tot januari blijven. \\
    \indent EN: He was going to stay until January. \\
\end{small}
%Sub-section: Polite Form
\textbf{Polite Form} \\
We also use zouden when we want to sound more polite. To politely express a
wish, we add the adverb \textit{graag (with pleasure,please)}, which give the
equivalent of the English \textit{I would like}.
\\
\begin{small}
    \indent NL: Ik zou graag een retourtje leiden willen. \\
    \indent EN: I would like a return ticket to leiden. \\
\end{small}
We also use \textit{zouden} to make a suggestion in the form, \textit{would
it not be better if}.
\\
\begin{small}
    \indent NL: Zou het niet makkelijker zijn als je geweoon een schaar gebruikte? \\
    \indent EN: Would it not be easier if you simply used scissors? \\
\end{small}
%Sub-section: Should
\textbf{Should} \\
Dutch does not have a verb for should, instead they use a combination of two
verbs, \textit{zouden moeten (should)}. If we place the infinitive
\textit{moeten(to have to)} after \textit{zouden} we get the English
equivalent of should.
\\
\begin{small}
    \indent NL: Dat zou je moten weten. \\
    \indent EN: You should know that. \\
\end{small}
When the required action is urgent or presented as a clear command, we often
simply use \textit{moeten}.
\\
\begin{small}
    \indent NL: Wat moeten we nu doen? \\
    \indent EN: What should we do not? \\
\end{small}
%------------------------------------------------------------------
% Section: Conditional Perfect
%------------------------------------------------------------------
\sectionhead{Conditional Perfect} \\
The conjugation is the same as the future perfect tense, with the only
difference being that the verb \textit{zullen} is conjugated in the simple past
tense.
\\
\begin{tabular}[t]{l l}
    \textbf{Conjugation} \\
    \hline
    \textit{Ik zou hebben/zijn} + \textit{p.p.} & \textit{We zouden hebben/zijn}
    + \textit{p.p.} \\
    \textit{Je zou hebben/zijn} + \textit{p.p.} & \textit{Jullie zouden
hebben/zijn} + \textit{p.p.} \\
    \textit{Hij zou hebben/zijn} + \textit{p.p.} & \textit{Ze zouden
hebben/zijn} + \textit{p.p.} \\
    \hline
\end{tabular}
\\ \\
We use the conditional perfect refers to hypothetical situations in the
past. Just like the \textit{simple conditional} tense, it mainly servers to
express a conditional clause.
\\
\begin{small}
    \indent NL: Als je goed had opgelet, zou hebben gemerkt dat \\
    \indent EN: If you had paid attention, you would have noticed that. \\
\end{small}
\textbf{Using the pluperfect to express a hypothetical situation in the past.}
\\
In Dutch we often use the \textit{pluperfect} tense to express a hypothetical
situation in the past.
\\
\begin{small}
    \indent NL: Ik had data zeker niet gedaan. \\
    \indent EN: I certainly would not have done that. \\
\end{small}
\textbf{Should have}
\\
For the past participle should have, the Dutch say \textit{hadden moeten +
infinitive}. The Dutch construction is int he pluperfect tense which we often
use instead of the conditional perfect.
\\
\begin{small}
     \indent NL: Ik had niet moeten zeggen.   \\
     \indent EN: I should not have said that. \\
\end{small}
%------------------------------------------------------------------
% Section: Aan het continuous
%------------------------------------------------------------------
\sectionhead{Aan het continuous}
\par{Just like in English, in Dutch we use the verb \textit{zijn (to be)} to
    express the continuous. The continuous is formed by a conjugation of
\textit{zijn} followed byt the words \textit{aan het} and the infinitive of the
verb that specifies the action.}
\\
\begin{tabular}[t]{l l l l}
    \textbf{Present tense} \\
    \hline
    \textit{Ik}     & \textit{ben}  & \textit{aan het} & \textit{infinitive} \\
    \textit{Je}     & \textit{bent} & \textit{aan het} & \textit{infinitive} \\
    \textit{Hij}    & \textit{is}   & \textit{aan het} & \textit{infinitive} \\
    \textit{We}     & \textit{zijn} & \textit{aan het} & \textit{infinitive} \\
    \textit{Jullie} & \textit{zijn} & \textit{aan het} & \textit{infinitive} \\
    \textit{Ze}     & \textit{zijn} & \textit{aan het} & \textit{infinitive} \\
    \hline
\end{tabular}
\begin{tabular}[t]{l l l l}
    \textbf{Past tense} \\
    \hline
    \textit{Ik}     & \textit{was}   & \textit{aan het} & \textit{infinitive} \\
    \textit{Je}     & \textit{was}   & \textit{aan het} & \textit{infinitive} \\
    \textit{Hij}    & \textit{was}   & \textit{aan het} & \textit{infinitive} \\
    \textit{We}     & \textit{waren} & \textit{aan het} & \textit{infinitive} \\
    \textit{Jullie} & \textit{waren} & \textit{aan het} & \textit{infinitive} \\
    \textit{Ze}     & \textit{zijn}  & \textit{aan het} & \textit{infinitive} \\
    \hline
\end{tabular}
\\
\par{You may not know the \textit{perfect} tense of \textit{aan het} continuous
does not perfectly follow the conjugation of \textit{zijn}. In fact, it does not
really follow the \textit{aan het} idea at all. This tense will be discussed
later.}
\\
\begin{tabular}[t]{l l}
    \textbf{Examples} \\
    \hline
    Ik ben de afwas aan het doen. & I am doing the dishes \\
    Mijn broer was zich aan het scheren. & My brother was shaving (himself) \\
    Det honden zijn aan het blaffen. & The dogs are barking. \\
    De kinderen waren aan het spelen. & The children were playing. \\
    \hline
\end{tabular}
\\
%-------------------------------------
% This section can be added in later
%-------------------------------------
% \sectionhead{Aan het perfect}
% \par{In Dutch the perfect \textit{aan het} continuous is not as commonly used as
% in English. For example in English we would say \textit{he has been working all
% day} in Dutch we would say \textit{he has worked all day}. So where in English
% we would use the \textit{perfect continuous}, the Dutch would instead opt for
% the \textit{pefect} tense.}
%
%------------------------------------------------------------------
% Section: Te continuous
%------------------------------------------------------------------
\\
\sectionhead{Te continuous}
\par{In Dutch, we have another way for forming the continuous. What we'll refer
    to here as the \textit{te-continuous} tense. We use this type of continuous
    to explress the physical situation of the subject. In this tense we use the
    verbs \textit{liggen (lie), zitten (sit), staan (stand), lopen (walk)} and
    less often \textit{hangen (to hang)} as auxiliary verbs.The
\textit{te-continuous} is formed by:}
\\
\indent \textit{Auxilary verb} + \textit{te} + \textit{infinitive} \\
\begin{tabular}[t]{l l}
    \textbf{Examples} \\
    \hline
    Hij light te slapen. & He is sleeping (while he is lying). \\
    Ik zit te lezen & I am reading (while he is sitting). \\
    Hij staat te koken. & He is cooking (while standing). \\
    Ik loop te zingen. & I am singing (while walking). \\
    Ze hangen maar te nietsen & They are not doing anything (while hanging
    around). \\
    Ik zit te denken.  & I am thinking (while sitting down). \\
    \hline
\end{tabular}
\par{Notice in the last example that the speaker mentions he is sitting down. If
you were asking this directly to someone it would not likely be necessary for
them to mention they were sitting. Yet in Dutch this information is often left
in regardless.}
\\
\textbf{Pefect tense of te-continuous}
\\
The \textit{perfect} tense of the \textit{te-continuous} is similar to the
\textit{pefect aan het-continuous}. We omit the preposition \textit{te (too)}
and we use an infinitive isntead of a \textit{past participle}. \\
\indent \textit{Hebben} + \textit{auxilary verb} + \textit{infinitive} \\
The \textit{pefect te continuous} is more common than the \textit{pefect aan het
continuous}. In Dutch however, we use the normal \textit{non-continuous} pefect
much more than the \textit{continuous perfect}. Notice the \textit{pefect tense}
does not get a past participle in the examples bellow.
\\
\begin{tabular}[t]{l l}
    \textbf{Examples} \\
    \hline
    Hij heeft liggen slapen. & He has been sleeping. \\
    Ik heb zitten lezen. & I have been reading. \\
    De was heeft hangen drogen. & The laundry has been drying. \\
    \hline
\end{tabular}
\\
%------------------------------------------------------------------
% Section: The passive voice
%------------------------------------------------------------------
\\
\sectionhead{Passive voice}
\par{In English we use verbs \textit{to be} or \textit{to become} for the
    passive. After teh passive, we can always use the grammar construction by
    adding the \textit{agent}. The passive is not a tense. A tense indicates the
    point in time an event occured. The passive voice can take all tenses, e.g.
the \textit{passive simple, passive perfect, pasive pluperfect, etc}}
\\
\begin{tabular}[t]{l l}
    \textbf{Active}  & \textbf{Passive} \\
    \hline
    I was your hair          & Your hair is washed (by me).         \\
    He has cleaned the house & The house has been cleaned (by him). \\
    The music excites me     & I am excited (by the music).         \\
    \hline
\end{tabular}
\\ \\
\textbf{Formation of the passive}
\\
\textit{Zijn (to be)} and \textit{worden (to become)} are irregular verbs. Just
like in English, we use the \textit{past participle} to form the
\textit{passive}. There is however, one important difference. Just compare the
English and Dutch phrases below.
\\
\begin{tabular}[t]{l l l l}
    \textbf{English}  & \textbf{Aux verb} & \textbf{Dutch} & \textbf{Aux verb}
    \\
    \hline
    My hair is washed       & to be & Mijn haar wordt gewassen & worden \\
    My hair was washed      & to be & Mijn haar werd gewassen  & worden \\
    My hair has been washed & to be & Mijn haar is gewassen    & zijn   \\
    My hair had been washed & to be & Mijn haar was gewassen   & zijn   \\
    \hline
\end{tabular}
\\
\par{In English, we use \textit{is} and \textit{was} for the \textit{simple}
    tense. Where as in Dutch, we use the exact same words for the pefect tense.
    In English, we use the same verb for both the simple and the perfect tenses
    \textit{to be}. In the samples above, we used \textit{to be}. In Dutch, on
the other hand, we use two differnt verbs: \textit{worden} and \textit{zijn}.}
\\
\textbf{Worden and zijn}
\par{\textit{Worden} is used if the event is stil lgoing on at this particular
moment. In English we would use the simple tense of \textit{to be, to become}
or, sometimes \textit{to get}. \textit{Zijn} is used when we want to express
that an event has alreadly occurred or been done \textit{pefect} tense. In
English, we would use the \textit{perfect} tense of \textit{to be}.}
\\
%------------------------------------------------------------------
% Section: The imperative
%------------------------------------------------------------------
\sectionhead{Imperative}
\par{For the \textit{imperative}, we use the \textit{present} tense of the 1st
person singular. For regular verbs, this is the verb stem.} 
\\
\begin{tabular}[t]{l l}
    \textbf{English}  & \textbf{Dutch} \\
    \hline
    Ge weg!             & Go away!              \\
    Gool data maar weg. & Just throw that away. \\
    Neem dat mee.       & Take that with you.   \\
    Kom eens mee.       & Just come here.       \\
    \hline
\end{tabular}
\\
\par{The verb \textit{zijn} has an irregular imperative. Instead of \textit{ben}
1st person singular, we use the stem of the verb \textit{wezen} an oldfashioned
form of \textit{zijn}.} 
\\
\begin{tabular}[t]{l l}
    \textbf{English}  & \textbf{Dutch} \\
    \hline
    Wees stil! & Be quiet! \\
    \hline
\end{tabular}
\\
\par{If you want to be polite, you can use the \textit{present} tense of the
more formal 2nd person. If you use the \textit{polite imperative}, you have to
add the personal pronoun u:} 
\\
\begin{tabular}[t]{l l}
    \textbf{English}  & \textbf{Dutch} \\
    \hline
    Vult u dit formulier in. & Fill in the form.   \\
    Betaalt u aan de kassa.  & Pay at the counter. \\
    Blijft u maar zitten     & Just remain seated. \\
    Weest u maar niet bang.  & Don't be afraid.    \\
    \hline
\end{tabular}
\\ \\
For the English expression \textit{let's...} we say: \textit{laten we...}
\\
\begin{tabular}[t]{l l}
    \textbf{English}           & \textbf{Dutch}        \\
    \hline
    Laten we wat gaan drinken. & Let's go for a drink. \\
    Laten we gaan.             & Let's go.             \\
    \hline
\end{tabular}
\\ \\
When a command is general and no particular person is addresed, we use the
infinitive. This is particularly common on public signs.
\\
\begin{tabular}[t]{l l}
    \textbf{English} & \textbf{Dutch}  \\
    \hline
    Niet roken.      & Do not smoke.   \\
    Niet aanraken    & Do not touch.   \\
    Trekken/Duwen    & Pull/Push       \\
    Niet storen      & Do not disturb. \\
    \hline
\end{tabular}
\\ \\
%------------------------------------------------------------------
% Section: Past Participle
%------------------------------------------------------------------
\sectionhead{Past Participle}
\begin{enumerate}[topsep=0pt, parsep=0pt, partopsep=0pt, itemsep=0pt]
    \item A past participle never ends in double \textit{d} or \textit{t}.
        \subitem \textit{Rusten} becomes \textit{Gerust}
        \subitem \textit{Bloeden} becomes \textit{Gebloed}
    \item All verbs that begin with the prefixes \textit{be-, er-, her-, ont-,}
        and \textit{ver-} do not get \textit{-ge} before the \textit{past
        participle}.
        \subitem \textit{Verdelen } becomes \textit{Verdeeld}
        \subitem \textit{Geschieden} becomes \textit{Geschied}
    \item In the category of \textit{her-} verbs, there are a few exceptions.
        \subitem \textit{Herbergen} becomes \textit{Geherberd}
        \subitem \textit{Herhuisvesten} becomes \textit{Geherstructureerd}
\end{enumerate}

\end{document}


% \begin{tabular}[t]{l l l}
%     \textbf{Infinitive} & \textbf{In English} & \textbf{Past Participle} \\
%     \hline
%     Verdelen & To distribute & Verdeeld \\
%     Geschieden & To happen & Geschied \\
%     Betalen & To pay & Betaald \\
%     Ontdekken & To discover & Ontdekt \\
%     Erkennen & To acknowledge & Erkend \\
%     Herkenner & To recognize & Herkend \\
% \end{tabular}

% \begin{tabular}[t]{l l l}
%     \textbf{Infinitive} & \textbf{In English} & \textbf{Past Participle} \\
%     \hline
%     Herbergen & To accommodate & Geherberd \\
%     Herhuisvesten & To relocate & Geherhuisvest \\
%     Herstructuren & To restructure & Geherstructureerd \\
% \end{tabular}

% when referring to an event that took place in the past we usually opt for the
% \textit{Present pefect}.
% \\
% For example using \textit{blaffen (to bark)}.

